---
title: "Stats 506, F18, Problem Set 3"
author: "Zi Wang, tlwangzi@umich.edu"
date: "11/21/2018"
output: html_document
---
## Question 1

```{r q1_source, echo = FALSE, message = FALSE}
library(ggplot2)
source('./ps3_q1.R')
```

### a.

```{r q1a_table, echo = FALSE}
cap = '**Table 1.** Percent of homes with stucco construction as 
  the major outside wall material within each division.'
p_stucco_tab = 
  p_stucco[order(-p_stucco) 
          ][,.(`Division` = division,
               `Precent of Stucco Homes (95% CI)` = 
                 sprintf('%4.1f%% (%4.1f, %4.1f)', p_stucco, lwr, upr))]
knitr::kable(p_stucco_tab, align = 'r', caption = cap)
```

```{r q1a_figure, echo = FALSE, fig.cap = cap}
cap = '**Figure 1.** Percent of homes with stucco construction as 
  the major outside wall material within each division.'
p_stucco = p_stucco[order(-p_stucco)
                   ][,`:=`(Division = 
                     factor(as.character(division), as.character(division)))]
ggplot(p_stucco, aes(x = Division, y = p_stucco) ) +
geom_col(fill = 'navy' ) +
geom_errorbar(aes(ymin = lwr, ymax = upr), 
              col = 'darkslategrey') +
theme_bw() +
ylab('Percent of Stucco Homes') +
xlab('Division') +
theme(axis.text.x = element_text(size = 8, angle = 90))
```

The Mountain South division has the highest proportion: 64%.   
The East South Central Division has the highest proportion: 0.4%.

### b.
```{r q1b_table1, echo = FALSE}
cap = '**Table 2.** Average total electricity usage in kilowatt hours 
in each division.'
m = qnorm(.975)
pwc = function(x) format(round(x), big.mark = ',')
kwh1 = kwh[order(-kwh)
         ][,.(`Division` = division,
              `Average total electricity usage in kilowatt hours(95% CI)` = 
              sprintf('%s, (%s - %s)', 
                     pwc(kwh), pwc(kwh - m*se_kwh), pwc(kwh + m*se_kwh)))]
knitr::kable(kwh1, align = 'r', caption = cap)
```

```{r q1b_figure1, echo = FALSE, fig.cap = cap}
cap = '**Figure 2.** Average total electricity usage in kilowatt hours 
in each division.'
kwh2 = kwh[order(-kwh)
          ][,`:=`(div = 
                  factor(as.character(division), as.character(division)))]
ggplot(kwh2, aes(x = div, y = kwh) ) +
geom_point() +
geom_errorbar(aes(ymin = lwr, ymax = upr)) +
coord_flip() +
theme_bw() +
ylab('kwh/home') +
xlab('')
```

```{r q1b_table2, echo = FALSE}
cap = '**Table 3.** Average total electricity usage in kilowatt hours 
in each division stratified by urban and rural status.'
kwh_div = kwh_div[,`:=`(div_avg = mean(kwh_div)), by = division
                 ][order(-div_avg)
                 ][,`:=`(div = factor(as.character(division), 
                         levels = unique(as.character(division))))]
kwh_div1 = 
  kwh_div[,.(`Division` = div,
             `Average Electricity Usage(95% CI)` = 
                sprintf('%s, (%6s - %6s)', 
                  pwc(kwh_div), 
                  pwc(kwh_div - m*se_kwh_div), pwc(kwh_div + m*se_kwh_div)),
             Rurality = ifelse(urban, 'Average kwh in Urban status(95% CI)', 
                                      'Average kwh in Rural status(95% CI)'))]
kwh_div1 = dcast(kwh_div1, `Division` ~ Rurality,
                 value.var = "Average Electricity Usage(95% CI)")
knitr::kable(kwh_div1, align  = 'r', cap = cap)
```

```{r q1b_figure2, echo = FALSE, fig.cap = cap}
cap = '**Figure 3.** Average total electricity usage in kilowatt hours 
in each division stratified by urban and rural status.'
kwh_div2 = kwh_div[,`:=`(Rurality = ifelse(urban, 'Urban', 'Rural'))]
ggplot(kwh_div2, aes(x = div, y = kwh_div, color = Rurality) ) +
geom_point(position = position_dodge(.5) ) +
geom_errorbar(aes(ymin = lwr, ymax = upr),
              position = position_dodge(.5)) +
scale_color_manual(values = c('navy', 'darkred')[2:1]) +
coord_flip() +
theme_bw() +
ylab('Average total electricity usage') +
xlab('')
```

### c.
```{r q1c_table, echo = FALSE}
cap = "**Table 4.** Disparity between urban and rural areas 
  in terms of the proportion of homes with internet access."
int_disp1 = int_disp[order(-est)
                    ][,.(`Census Division` = division, 
                         `Urban Internet Access, % (95% CI)` = Urban,
                         `Rural Internet Access, % (95% CI)` = Rural, 
                         `Difference, % (95% CI)` = Diff)]
knitr::kable(int_disp1, align = 'r', caption = cap )
```

The Mountain South Division has the largest disparity 
between urban and rural areas in terms of the proportion of homes 
with internet access: 18.5%  

```{r q1c_graph, echo = FALSE, fig.cap = cap}
cap = "**Figure 4.** Disparity between urban and rural areas 
  in terms of the proportion of homes with internet access."
int_div2 = int_div[,`:=`(Rurality = ifelse(urban, 'Urban', 'Rural'),
                         division = factor(as.character(division),
                           as.character(
                           (int_disp[order((Rural))]$division))))] 
ggplot(int_div2, aes(x = division, y = int_div, fill = Rurality) ) +
geom_col(position = position_dodge() ) +
geom_errorbar(aes(ymin = lwr, ymax = upr), 
              position = position_dodge(),
              col = 'slategrey') +
theme_bw() + 
xlab('') +
ylab('Proportion of homes with internet access') +
ylim(c(0, 100)) +
coord_flip() +
scale_fill_manual(values = c('darkred', 'navy'))
```

### d.

Question: What percent of homes have rented(KOWNRENT == 2) within each division? 
Which division has the highest proportion? Which has the lowest?

```{r q1d_table, echo = FALSE, fig.height = 4, fig.cap = cap}
cap = '**Table 5.** Percent of homes rented(KOWNRENT == 2) 
within each division.'
p_rent_tab = 
  p_rent[order(-p_rent) 
        ][,.(`Division` = division,
             `Precent of Stucco Homes (95% CI)` = 
               sprintf('%4.1f%% (%4.1f, %4.1f)', p_rent, lwr, upr))]
knitr::kable(p_rent_tab, align = 'r', caption = cap)
```

```{r q1d_figure, echo = FALSE, fig.cap = cap}
cap = '**Figure 5.** Percent of homes rented(KOWNRENT == 2) 
within each division.'
p_rent = p_rent[order(-p_rent)
               ][,`:=`(Division = 
                 factor(as.character(division), as.character(division)))]
ggplot(p_rent, aes(x = Division, y = p_rent) ) +
geom_col(fill = 'navy' ) +
geom_errorbar(aes(ymin = lwr, ymax = upr), 
              col = 'darkslategrey') +
theme_bw() +
ylab('Percent of homes rented') +
xlab('Division') +
theme(axis.text.x = element_text(size = 8, angle = 90))
```

The Pacific division has the highest proportion: 43.7%.   
The East South Central Division has the lowest proportion: 30.1%.

\pagebreak

## Question 2 

```{r q2_source, echo = FALSE, message = FALSE}
source('./ps3_q2.R')
```

### a.

The function:
```{r q2a_function, echo = TRUE, eval = FALSE}
MC_analysis = function(X, beta, sigma, mc_rep){
  result = matrix(0, nrow = p, ncol = mc_rep)
  var = rep(0,p)
  QR = qr(t(X) %*% X)
  R = qr.R(QR)
  Q = qr.Q(QR)
  for (i in 1:mc_rep){
    Y = X %*% beta 
    Y1 = rnorm(n, mean = 0, sd = sigma)
    dim(Y1) = c(n,1)
    Y = Y + Y1
    # i.
    beta_hat = solve(R, t(Q) %*% t(X) %*% Y)
    ## ii.
    Y_hat = X %*% beta_hat
    sigmasq_hat = sum((Y-Y_hat)^2)/(n-p)
    ## iii.
    v = (sigmasq_hat) * solve(R) %*% t(Q)
    var = diag(v)
    ## iv.
    Z = beta_hat/sqrt(var)
    result[,i] = 2*(1-pnorm(abs(Z)))
  }
  result
}
```

Test the function, show the maximum of the finite precision error:
```{r q2a, echo = TRUE}
# Test the function: MC_analysis()
n = 1000
p = 100
mc_rep = 1
sigma = 1
beta = rep(0,p)
for (i in 1:10) beta[i] = sigma^2
X = rnorm(n*p)
dim(X) = c(n,p)
set.seed(42)
test = MC_analysis(X, beta, sigma, mc_rep)
set.seed(42)
Y = X%*%beta + rnorm(n, mean = 0, sd = sigma)
fit0 = lm(Y ~ 0 + X)
p_value = summary(fit0)$coefficients[,'Pr(>|t|)']
ans = max(abs(test[,1]-p_value))
ans
```


### b.
```{r q2b, echo = TRUE, eval = FALSE}
n = 1000
p = 100
sigma = 30                   # sigma
S = matrix(0, nrow = p, ncol = p)
for (i in 1:p)
  S[i,i] = runif(1, min = 1, max = p)
R = chol(S)
X = rnorm(n*p)
dim(X) = c(n,p)
X = X %*% R                  # X
beta = rep(0,p)
for (i in 1:10) beta[i] = 1  # beta
mc_rep = 1000                # mc_rep
result = MC_analysis(X, beta, sigma, mc_rep)
```

### c.
The evaluate() function:
```{r q2c, echo = TRUE, eval = FALSE}
evaluate = function(result, indice){
  alpha = 0.05
  m = qnorm(.975)
  p = nrow(result)
  mc_rep = ncol(result)
  FWER = rep(0,mc_rep)
  FDR = rep(0,mc_rep)
  Sensitivity = rep(0,mc_rep)
  Specificity = rep(0,mc_rep)
  for (j in 1:mc_rep)
  {
    tp = 0  # true positive
    tn = 0  # true negative
    fp = 0  # false positive
    fn = 0  # false negative
    fn = sum(result[indice, j] >= alpha)  # H0 is False, and accept H0
    tp = sum(result[indice, j] < alpha)   # H0 is false, and reject H0
    tn = sum(result[-indice, j] >= alpha) # H0 is true, and accept H0
    fp = sum(result[-indice, j] < alpha)  # H0 is true, and reject H0
    FWER[j] = (fp>=1)
    FDR[j] = fp/(fp+tp)
    Sensitivity[j] = tp/(tp+fn)
    Specificity[j] = tn/(tn+fp)
  }
  res = tibble(quan = c('The family wise error rate',
                        'The false discovery rate',
                        'The sensitivity',
                        'The specificity'),
               est = c(mean(FWER), mean(FDR), 
                       mean(Sensitivity), mean(Specificity)),
               se =  c(sd(FWER)/sqrt(mc_rep), 
                       sd(FDR)/sqrt(mc_rep), 
                       sd(Sensitivity)/sqrt(mc_rep), 
                       sd(Specificity)/sqrt(mc_rep)))
  res = res %>%
    group_by(quan) %>%
    mutate(lwr = pmax(est - m*se, 0), upr = pmin(est + m*se, 1),
           ci = sprintf('%4.3f (%4.3f, %4.3f)', est, lwr, upr)) %>%
    ungroup %>%
    transmute(Quantities = quan, ci)
  res
}
```

### d.
```{r q2d, echo = TRUE, eval = FALSE}
alpha = 0.05 # Confidence level
indice = vector()
for (i in 1:p)
  if (beta[i]!=0) indice = cbind(indice,i)
fit1 = evaluate(result,indice) %>%
  transmute(Quantities, "Uncorrected(95% CI)" = ci)
result_adj_bon = apply(result, 2,function(x) p.adjust(x,method = 'bonferroni'))
fit2 = evaluate(result_adj_bon,indice) %>%
  transmute(Quantities, "Bonferroni(95% CI)" = ci)
result_adj_holm = apply(result, 2, function(x) p.adjust(x,method = 'holm'))
fit3 = evaluate(result_adj_holm,indice) %>%
  transmute(Quantities, "Holm(95% CI)" = ci)
result_adj_bh = apply(result, 2, function(x) p.adjust(x,method = 'BH'))
fit4 = evaluate(result_adj_bh,indice) %>%
  transmute(Quantities, "BH Correction(95% CI)" = ci)
result_adj_by = apply(result, 2, function(x) p.adjust(x,method = 'BY'))
fit5 = evaluate(result_adj_by,indice) %>%
  transmute(Quantities, "BY Correction(95% CI)" = ci)
```

### e.
```{r q2e, echo = TRUE, fig.height = 4, fig.cap = cap}
fit = fit1 %>%
  left_join(fit2, by = "Quantities") %>%
  left_join(fit3, by = "Quantities") %>%
  left_join(fit4, by = "Quantities") %>%
  left_join(fit5, by = "Quantities")
cap = '**Table 6.** The  Monte Carlo estimates of the four quantities 
in the uncorrected and other four corrected situations'
knitr::kable(fit, align  = 'r', cap = cap)
```

Discussion:  
1. Both 4 corrections give smaller value of the family wise error rate than 
that of the uncorrected. Besides, the Bonferroni correction and Holm correction 
give small values of the family wise error rate compared to results of 
the uncorrected because they are designed to give strong control of 
the family wise error rate.    
2. Both 4 corrections give smaller value of the false discovery rate than 
that of the uncorrected. Besides, the BH correction and BY correction give 
small values of the false discovery rate compared to results of 
the uncorrected because they are designed to give control of 
the false discovery rate.  
3. The sensitivity are both very large among the 4 corrected methods 
and the uncorrected.     
4. The specificity are both very large among the 4 corrected methods 
and the uncorrected.    

\pagebreak

## Question 3

```{r q3_source, echo = FALSE, message = FALSE}
source('./ps3_q3.R')
```

### a.
```{r q3a, echo = TRUE, eval = FALSE}
# Import data and select key items
df = fread('mtcars.csv')
df = df[,.(mpg,cyl,disp,hp,wt)]

# Center disp, hp, and wt within cyl and then computer regression coefficients.
beta_cyl = df[order(cyl)
             ][,`:=`(disp_gc = disp - mean(disp),
                     hp_gc = hp - mean(hp),
                     wt_gc = wt - mean(wt)), by = cyl
             ][,.(dispXmpg = sum(mpg*disp_gc), disp_sq = sum(disp_gc*disp_gc),
                  hpXmpg = sum(mpg*hp_gc), hp_sq = sum(hp_gc*hp_gc),
                  wtXmpg = sum(mpg*wt_gc), wt_sq = sum(wt_gc*wt_gc)), by = cyl
             ][,`:=`(beta_cyl_disp = dispXmpg / (disp_sq),
                     beta_cyl_hp = hpXmpg / (hp_sq),
                     beta_cyl_wt = wtXmpg / (wt_sq))
             ][,.(cyl, beta_cyl_disp, beta_cyl_hp, beta_cyl_wt)]
beta_cyl

# Export data
fwrite(beta_cyl,file='mpg_betas_by_cyl.csv')
```

Results:
```{r q3a_result, echo = FALSE}
beta_cyl
```

### b.
The function:
```{r q3b_function, echo = TRUE, eval = FALSE}
compute_reg_coef = function(df, response, predictor, group){
# Center disp, hp, and wt within cyl and then computer regression coefficients.
beta_cyl = df[,.(get(response),get(predictor),get(group))
             ][,.(response = V1, predictor = V2, group = V3)
             ][order(group)
             ][,`:=`(predictor_gc = predictor - mean(predictor)), by = group
             ][,.(preXres = sum(response * predictor_gc),
                  predictor_sq = sum(predictor_gc * predictor_gc)), by = group
             ][,.(group, beta_group_predictor = preXres / predictor_sq)]
beta_cyl
}
```

Test the function:
```{r q3b_test, echo = TRUE}
# beta_cyl_disp
beta1 = compute_reg_coef(df, "mpg", "disp", "cyl")

# beta_cyl_hp
beta2 = compute_reg_coef(df, "mpg", "hp", "cyl")

# beta_cyl_wt
beta3 = compute_reg_coef(df, "mpg", "wt", "cyl")

# Compare the difference between results in (a)
c(max(beta1[,2] - beta_cyl[,2]),
  max(beta2[,2] - beta_cyl[,3]),
  max(beta3[,2] - beta_cyl[,4]))
```



### c.
```{r q3c, echo = TRUE}
# Center disp, hp, and wt within cyl and then computer regression coefficients.
beta_cyl_dplyr = df %>%
  group_by(cyl) %>%
  mutate(disp_gc = disp - mean(disp),
         hp_gc = hp - mean(hp),
         wt_gc = wt - mean(wt),
         dispXmpg = mpg*disp_gc, disp_sq = disp_gc*disp_gc,
         hpXmpg = mpg*hp_gc, hp_sq = hp_gc*hp_gc,
         wtXmpg = mpg*wt_gc, wt_sq = wt_gc*wt_gc) %>%
  summarize_at(vars(dispXmpg:wt_sq), sum) %>%
  mutate(beta_cyl_disp = dispXmpg / (disp_sq),
         beta_cyl_hp = hpXmpg / (hp_sq),
         beta_cyl_wt = wtXmpg / (wt_sq)) %>%
  select(cyl, beta_cyl_disp, beta_cyl_hp, beta_cyl_wt)
```

Results:
```{r q3c_result, echo = FALSE}
beta_cyl_dplyr
```

### d.
The function:
```{r q3d_function, echo = TRUE}
compute_reg_coef_dplyr = function(df, response, predictor, group){
  # quote the variable with quo(), then unquote it in the dplyr call with !!
  response = enquo(response)
  predictor = enquo(predictor)
  group = enquo(group)
  
  # Use quo_name() to convert the input expression to a string,
  #   so we create the new names by pasting together strings
  predictor_gc = paste0(quo_name(predictor),"_gc")
  preXres = paste0(quo_name(predictor), "X", quo_name(response))
  predictor_sq = paste0(quo_name(predictor),"_sq")
  beta_group_predictor = 
    paste0("beta_",quo_name(group),"_",quo_name(predictor))
  beta_cyl = df %>%
    group_by(!!group) %>%
    
    # !! mean_name = mean(!! expr) is not valid R code, 
    #   so we need to use the := helper 
    # Use quo_name() to convert the input expression to a string,
    #   and then use .data[[]] to use the column we want
    mutate(!!predictor_gc := !!predictor - mean(!!predictor),
           !!preXres := sum(!!response * .data[[quo_name(predictor_gc)]]),
           !!predictor_sq := sum(.data[[quo_name(predictor_gc)]] *
                                 .data[[quo_name(predictor_gc)]])) %>%
    mutate(!!beta_group_predictor := .data[[quo_name(preXres)]] / 
                                     .data[[quo_name(predictor_sq)]]) %>%
    arrange(!!group) %>%
    ungroup %>%
    distinct(!!group,.data[[!!beta_group_predictor]])
  beta_cyl
}
```

Test the function:
```{r q3d_test, echo = TRUE}
# beta_cyl_disp
beta1_dplyr = compute_reg_coef_dplyr(df, mpg, disp, cyl)

# beta_cyl_hp
beta2_dplyr = compute_reg_coef_dplyr(df, mpg, hp, cyl)

# beta_cyl_wt
beta3_dplyr = compute_reg_coef_dplyr(df, mpg, wt, cyl)

# Compare the difference between results in (a)
c(max(beta1_dplyr$beta_cyl_disp - beta_cyl[,2]),
  max(beta2_dplyr$beta_cyl_hp - beta_cyl[,3]),
  max(beta3_dplyr$beta_cyl_wt - beta_cyl[,4]))
```