---
title: "Stats 506, F18, Problem Set 2"
author: "Zi Wang, tlwangzi@umich.edu"
date: "10/19/2018"
output: pdf_document
---
## Question 1 

```{r setup_q1, echo = FALSE, include = FALSE}
suppressMessages(library("tidyverse"))
recs_tib = readr::read_delim("./recs2015_usage.csv",delim = ",")
recs_2015 = recs_tib %>%
  mutate(lwr = Estimate-qnorm(.975)*SE,upr = Estimate+qnorm(.975)*SE) %>%
  select(Kind,Estimate,lwr,upr)
```

```{r print_q1, echo = FALSE, fig.height = 4, fig.cap = cap}
cap = "Estimates and 95% confidence intervals of the four national means
       for residential energy consumption"
cols = c('Total','Estimate','lwr of 95% CI','upr of 95% CI')
knitr::kable(recs_2015,caption = cap,col.names = cols)
```

\pagebreak

## Question 2

### a.

Use "import" command to read both data sets into Stata
Use "merge" command to merge them together by the participant id SEQN.

Stata code:  
import sasxport "https://wwwn.cdc.gov/Nchs/Nhanes/2005-2006/OHX_D.XPT"  
save oral, replace  
import sasxport "https://wwwn.cdc.gov/Nchs/Nhanes/2005-2006/DEMO_D.XPT"  
merge 1:1 seqn using oral  
save syn, replace  

### b.

The summary of the model:
\begin{tabular}{lc} \hline
 & (1) \\
VARIABLES & Model sign~ridagemn \\ \hline
 &  \\
ridagemn & 0.0697*** \\
 & (0.00257) \\
Constant & -8.359*** \\
 & (0.323) \\
 &  \\
 Observations & 7,563 \\ \hline
\multicolumn{2}{c}{ Standard errors in parentheses} \\
\multicolumn{2}{c}{ *** p$<$0.01, ** p$<$0.05, * p$<$0.1} \\
\end{tabular}

The BIC of the model: 1533.407

The age at which 25% of individuals lose their primary upper right 2nd bicuspid: 104

The age at which 50% of individuals lose their primary upper right 2nd bicuspid: 120

The age at which 75% of individuals lose their primary upper right 2nd bicuspid: 136

The range of representative age values: 8, 9, 10, 11, 12

### c.
Add gender to the model and show the summary:
\begin{tabular}{lc} \hline
 & (1) \\
VARIABLES & Model ridagemn~riagendr \\ \hline
 &  \\
ridagemn & 0.0697*** \\
 & (0.00257) \\
2.riagendr & 0.0702 \\
 & (0.132) \\
Constant & -8.397*** \\
 & (0.332) \\
 &  \\
 Observations & 7,563 \\ \hline
\multicolumn{2}{c}{ Standard errors in parentheses} \\
\multicolumn{2}{c}{ *** p$<$0.01, ** p$<$0.05, * p$<$0.1} \\
\end{tabular}

The BIC: 1542.055  
It is higher than 1533.407, so do not retain gender.


Add category Mexican American to the model and show the summary:
\begin{tabular}{lc} \hline
 & (1) \\
VARIABLES & Model sign~ridagemn+mex \\ \hline
 &  \\
ridagemn & 0.0704*** \\
 & (0.00265) \\
1.mex & 0.0386 \\
 & (0.143) \\
Constant & -8.459*** \\
 & (0.337) \\
 &  \\
 Observations & 7,246 \\ \hline
\multicolumn{2}{c}{ Standard errors in parentheses} \\
\multicolumn{2}{c}{ *** p$<$0.01, ** p$<$0.05, * p$<$0.1} \\
\end{tabular}

The BIC: 1542.285  
It is higher than 1533.407, so do not retain category Mexican American.


Add category Non-Hispanic Black to the model and show the summary:
\begin{tabular}{lc} \hline
 & (1) \\
VARIABLES & Model sign~ridagemn+black \\ \hline
 &  \\
ridagemn & 0.0701*** \\
 & (0.00259) \\
1.black & 0.521*** \\
 & (0.145) \\
Constant & -8.567*** \\
 & (0.334) \\
 &  \\
 Observations & 7,563 \\ \hline
\multicolumn{2}{c}{ Standard errors in parentheses} \\
\multicolumn{2}{c}{ *** p$<$0.01, ** p$<$0.05, * p$<$0.1} \\
\end{tabular}

The BIC: 1529.281  
It is lower than 1533.407, so retain category Non-Hispanic Black.


Add category Other to the model and show the summary:
\begin{tabular}{lc} \hline
 & (1) \\
VARIABLES & Model sign~ridagemn+black+other \\ \hline
 &  \\
ridagemn & 0.0703*** \\
 & (0.00260) \\
1.black & 0.567*** \\
 & (0.149) \\
1.other & 0.337 \\
 & (0.233) \\
Constant & -8.636*** \\
 & (0.339) \\
 &  \\
 Observations & 7,563 \\ \hline
\multicolumn{2}{c}{ Standard errors in parentheses} \\
\multicolumn{2}{c}{ *** p$<$0.01, ** p$<$0.05, * p$<$0.1} \\
\end{tabular}

The BIC: 1536.103  
It is higher than 1529.281, so do not retain category Other.

Add poverty income ratio to the model:
\begin{tabular}{lc} \hline
 & (1) \\
VARIABLES & Model sign~ridagemn+black+indfmpir \\ \hline
 &  \\
ridagemn & 0.0714*** \\
 & (0.00271) \\
1.black & 0.495*** \\
 & (0.149) \\
indfmpir & -0.119*** \\
 & (0.0454) \\
Constant & -8.460*** \\
 & (0.351) \\
 &  \\
 Observations & 7,246 \\ \hline
\multicolumn{2}{c}{ Standard errors in parentheses} \\
\multicolumn{2}{c}{ *** p$<$0.01, ** p$<$0.05, * p$<$0.1} \\
\end{tabular}

The BIC: 1462.895  
It is lower than 1529.281, so retain poverty income ratio.


The final model:
\begin{tabular}{lc} \hline
 & (1) \\
VARIABLES & Model sign~ridagemn+black+indfmpir \\ \hline
 &  \\
ridagemn & 0.0714*** \\
 & (0.00271) \\
1.black & 0.495*** \\
 & (0.149) \\
indfmpir & -0.119*** \\
 & (0.0454) \\
Constant & -8.460*** \\
 & (0.351) \\
 &  \\
 Observations & 7,246 \\ \hline
\multicolumn{2}{c}{ Standard errors in parentheses} \\
\multicolumn{2}{c}{ *** p$<$0.01, ** p$<$0.05, * p$<$0.1} \\
\end{tabular}


### d.
#### 1.
Adjusted Predictions at the mean (for other values) at the representative ages
(From 8 to 12):
\begin{tabular}{lc} \hline
 & (1) \\
VARIABLES & Adjusted predictions \\ \hline
 &  \\
1bn.\_at & 0.146*** \\
 & (0.0128) \\
2.\_at & 0.287*** \\
 & (0.0167) \\
3.\_at & 0.486*** \\
 & (0.0174) \\
4.\_at & 0.690*** \\
 & (0.0155) \\
5.\_at & 0.840*** \\
 & (0.0118) \\
 &  \\
 Observations & 7,246 \\ \hline
\multicolumn{2}{c}{ Standard errors in parentheses} \\
\multicolumn{2}{c}{ *** p$<$0.01, ** p$<$0.05, * p$<$0.1} \\
\end{tabular}

The plot can be seen from the corresponding "marginsplot" command in ps2_q2.do

#### 2.
The marginal effects at the mean of black 
at the same representative ages:
\begin{tabular}{lcc} \hline
 & (1) & (2) \\
VARIABLES & The marginal effects at the mean & The marginal effects at the mean \\ \hline
 &  &  \\
1bno.\_at & 0 &  \\
 & (0) &  \\
2o.\_at & 0 &  \\
 & (0) &  \\
3o.\_at & 0 &  \\
 & (0) &  \\
4o.\_at & 0 &  \\
 & (0) &  \\
5o.\_at & 0 &  \\
 & (0) &  \\
1bn.\_at &  & 0.0668*** \\
 &  & (0.0217) \\
2.\_at &  & 0.106*** \\
 &  & (0.0328) \\
3.\_at &  & 0.123*** \\
 &  & (0.0365) \\
4.\_at &  & 0.101*** \\
 &  & (0.0290) \\
5.\_at &  & 0.0616*** \\
 &  & (0.0175) \\
 &  &  \\
 Observations & 7,246 & 7,246 \\ \hline
\multicolumn{3}{c}{ Standard errors in parentheses} \\
\multicolumn{3}{c}{ *** p$<$0.01, ** p$<$0.05, * p$<$0.1} \\
\end{tabular}

The plot can be seen from the corresponding "marginsplot" command in ps2_q2.do

#### 3.
Average Marginal Effect of black at the representative ages:
\begin{tabular}{lcc} \hline
 & (1) & (2) \\
VARIABLES & Average marginal effect & Average marginal effect \\ \hline
 &  &  \\
1bno.\_at & 0 &  \\
 & (0) &  \\
2o.\_at & 0 &  \\
 & (0) &  \\
3o.\_at & 0 &  \\
 & (0) &  \\
4o.\_at & 0 &  \\
 & (0) &  \\
5o.\_at & 0 &  \\
 & (0) &  \\
1bn.\_at &  & 0.0671*** \\
 &  & (0.0217) \\
2.\_at &  & 0.105*** \\
 &  & (0.0326) \\
3.\_at &  & 0.122*** \\
 &  & (0.0363) \\
4.\_at &  & 0.100*** \\
 &  & (0.0289) \\
5.\_at &  & 0.0619*** \\
 &  & (0.0176) \\
 &  &  \\
 Observations & 7,246 & 7,246 \\ \hline
\multicolumn{3}{c}{ Standard errors in parentheses} \\
\multicolumn{3}{c}{ *** p$<$0.01, ** p$<$0.05, * p$<$0.1} \\
\end{tabular}

The plot can be seen from the corresponding "marginsplot" command in ps2_q2.do


### e.
Compare the two model:
\begin{tabular}{lcc} \hline
 & (1) & (2) \\
VARIABLES & Model sign~ridagemn+black+indfmpir & Model svy:sign~ridagemn+black+indfmpir \\ \hline
 &  &  \\
ridagemn & 0.0714*** & 0.0619*** \\
 & (0.00271) & (0.00723) \\
1.black & 0.495*** &  \\
 & (0.149) &  \\
indfmpir & -0.119*** & -0.0812 \\
 & (0.0454) & (0.0522) \\
black &  & 0.543*** \\
 &  & (0.146) \\
Constant & -8.460*** & -7.516*** \\
 & (0.351) & (0.862) \\
 &  &  \\
 Observations & 7,246 & 7,246 \\ \hline
\multicolumn{3}{c}{ Standard errors in parentheses} \\
\multicolumn{3}{c}{ *** p$<$0.01, ** p$<$0.05, * p$<$0.1} \\
\end{tabular}

Comments on the differences:

The coefficient of the variable ridagemn is more signficant from the model using svy than the other one.

The coefficient of the variable black is less signficant from the model using svy than the other one.

The coefficient of the variable indfmpir is more signficant from the model using svy than the other one.

The coefficient of the constant is more signficant from the model using svy than the other one.


Reason: the coefficient with lower p-value is more significant.

\pagebreak

## Question 3

```{r setup_q3, echo = FALSE, include = FALSE}
suppressMessages(library("tidyverse"))
library(margins)
library(prediction)
library(foreign)
source('./ps2_q3.R')
```

### a.

```{r q3a, eval = FALSE}
health = read.xport("OHX_D.XPT")
demo = read.xport("DEMO_D.XPT")
syn = demo %>%
  left_join(health,by = 'SEQN')
```

### b.
The original model:
```{r q3b, eval = FALSE}
log_reg = glm(sign ~ RIDAGEMN,family = binomial(link = 'logit'),
              data = syn_select)
```

The summary of the model:
```{r q3b_sum}
summary(log_reg)
```

The BIC of the model:
```{r q3b_BIC,echo = FALSE}
BIC_orig
```

The age at which 25% of individuals lose their primary upper right 2nd bicuspid:
```{r q3b_25, echo = FALSE} 
age_25
```

The age at which 50% of individuals lose their primary upper right 2nd bicuspid:
```{r q3b_50, echo = FALSE} 
age_50
```

The age at which 75% of individuals lose their primary upper right 2nd bicuspid:
```{r q3b_75, echo = FALSE} 
age_75
```

The lower bound and upper bound of the range of representative age values:
```{r q3b_range, echo = FALSE}
c(year_25,year_75)
```


### c.
Add gender to the model and show the summary:
```{r q3c_gen, eval = FALSE}
log_reg_gen = glm(sign ~ RIDAGEMN + RIAGENDR,
                  family = binomial(link = 'logit'),data = syn_select)
```

```{r q3c_gen_sum}
summary(log_reg_gen)
```

the BIC:
```{r q3c_BIC_gen, echo = FALSE}
BIC_gen
```
 
It is higher than BIC_orig = 1533.407, so do not retain gender.


Add category Mexican American to the model and show the summary:
```{r q3c_W_Mex, eval = FALSE}
log_reg_W_Mex = glm(sign ~ RIDAGEMN + Mex,
                    family = binomial(link = 'logit'),data = syn_select_1)
```

```{r q3c_W_Mex_sum}
summary(log_reg_W_Mex)
```

the BIC:
```{r q3c_BIC_W_Mex, echo = FALSE}
BIC_W_Mex
```
 
It is higher than BIC_orig = 1533.407, so do not retain category Mexican American.


Add category Non-Hispanic Black to the model and show the summary:
```{r q3c_W_Black, eval = FALSE}
log_reg_W_Black = glm(sign ~ RIDAGEMN + Black,
                      family = binomial(link = 'logit'),data = syn_select_1)
```

```{r q3c_W_Black_sum}
summary(log_reg_W_Black)
```

the BIC:
```{r q3c_BIC_W_Black}
BIC_W_Black
```
 
It is lower than BIC_orig = 1533.407, which means it improves the BIC, 
so retain category Non-Hispanic Black.


Add category Other to the model and show the summary:
```{r q3c_W_Black_Other, eval = FALSE}
log_reg_W_Black_Other = glm(sign ~ RIDAGEMN + Black + Other,
                            family = binomial(link = 'logit'),
                            data = syn_select_1)
```

```{r q3c_W_Black_Other_sum}
summary(log_reg_W_Black_Other)
```

the BIC:
```{r q3c_BIC_W_Black_Other}
BIC_W_Black_Other
```
 
It is higher than BIC_W_Black = 1529.281, so do not retain category Other.

Add poverty income ratio(PIR) to the model: 
```{r q3c_W_Black_PIR, eval = FALSE}
log_reg_W_Black_PIR = glm(sign ~ RIDAGEMN + Black + INDFMPIR,
                          family = binomial(link = 'logit'),
                          data = syn_select_2)
```

```{r q3c_W_Black_PIR_sum}
summary(log_reg_W_Black_PIR)
```

the BIC:
```{r q3c_BIC_W_Black_PIR}
BIC_W_Black_PIR
```

It is lower than BIC_W_Black = 1529.281, which means it improves the BIC, 
so retain category poverty income ratio.

The final model: sign ~ RIDAGEMN + Black + INDFMPIR
```{r q3c_Final, echo = FALSE}
summary(log_reg_W_Black_PIR)
```


### d.
#### 1.
Adjusted Predictions at the mean (for other values) at the representative ages
(From 8 to 12):

```{r q3d_apm_table, echo =FALSE, fig.height = 4, fig.cap = cap}
syn_select_apm_1 = syn_select_apm %>%
  mutate(fit=adp$fit, se=adp$se.fit, lwr=fit - 2*se, upr=fit + 2*se,
         age = RIDAGEMN/12)
syn_apm_tab = syn_select_apm_1 %>%
  select(age, fit)
cap = "Adjusted predictions at the mean (for other values) 
           at each of the representative age"
cols=c('Age','Adjusted predictions at the mean')
knitr::kable(syn_apm_tab,caption = cap,col.names = cols)
```

```{r q3d_apm_plot, echo =FALSE}
syn_select_apm %>%
  mutate(fit=adp$fit, se=adp$se.fit, lwr=fit - 2*se, upr=fit + 2*se,
         age = RIDAGEMN/12) %>%
  mutate(age=factor(age,age[order(fit)])) %>% # order
  ggplot(aes(y=age, x=fit)) +
  geom_point() + 
  geom_errorbarh(aes(xmin=lwr, xmax=upr), height=.1) + 
  geom_vline(aes(xintercept = fit), lty = 'dashed', color='grey', alpha =.5) + 
  xlab('Adjusted predictions at the mean') +
  ggtitle('Adjusted predictions at the mean (for other values) 
           at each of the representative ages') +
  theme_bw() +
  xlim(c(0,1))
```

#### 2.
The marginal effects at the mean of Black at the same representative ages:

```{r q3d_mem_table, echo =FALSE, fig.height = 4, fig.cap = cap}
syn_select_mem_1 = syn_select_mem %>%  
  distinct(RIDAGEMN,INDFMPIR) %>%
  mutate(fit=amm[,1], se=amm[,2], lwr=fit - 2*se, upr=fit + 2*se,
         age = RIDAGEMN/12)
syn_mem_tab = syn_select_mem_1 %>%
  select(age, fit)
cap = "The marginal effects at the mean of categorical variables 
           ridrethBlack at the same representative age"
cols=c('Age','The marginal effects at the mean')
knitr::kable(syn_mem_tab,caption = cap,col.names = cols)
```

```{r q3d_mem_plot, echo = FALSE}
syn_select_mem %>% 
  distinct(RIDAGEMN,INDFMPIR) %>%
  mutate(fit=amm[,1], se=amm[,2], lwr=fit - 2*se, upr=fit + 2*se,
         age = RIDAGEMN/12) %>%
  mutate(age=factor(age,age[order(fit)])) %>% # order
  ggplot(aes(y=age, x=fit)) +
  geom_point() + 
  geom_errorbarh(aes(xmin=lwr, xmax=upr), height=.1) + 
  geom_vline( aes( xintercept = fit), lty = 'dashed', color='grey', alpha =.5 ) + 
  xlab('The marginal effects at the mean') +
  ggtitle('The marginal effects at the mean of categorical variables 
           ridrethBlack at the same representative ages') +
  theme_bw() +
  xlim(c(0,0.30))
```

#### 3.
Use "margins" command in the "margins" package with the "at" option:
```{r q3d_ame}
q_tot = c(96,108,120,132,144)
mem0 = margins(log_reg_W_Black_PIR_1, at = list(RIDAGEMN = q_tot))
mem0
```

So, Average Marginal Effect of Black at the representative ages:

```{r q3d_ame_table, echo =FALSE, fig.height = 4, fig.cap = cap}
syn_select_ame_1 = syn_select_ame %>%  
  mutate(fit=Black, se=ame[,2], lwr=fit - 2*se, upr=fit + 2*se,
         age = RIDAGEMN/12)
syn_ame_tab = syn_select_ame_1 %>%
  select(age, fit)
cap = "The average marginal effect of Black at the representative ages"
cols=c('Age','The average marginal effect')
knitr::kable(syn_ame_tab,caption = cap,col.names = cols)
```

```{r q3d_ame_plot, echo = FALSE}
syn_select_ame %>% 
  mutate(fit=Black, se=ame[,2], lwr=fit - 2*se, upr=fit + 2*se,
         age = RIDAGEMN/12) %>%
  mutate(age=factor(age,age[order(fit)])) %>% # order
  ggplot(aes(y=age, x=fit)) +
  geom_point() + 
  geom_errorbarh(aes(xmin=lwr, xmax=upr), height=.1) + 
  geom_vline( aes( xintercept = fit), lty = 'dashed', color='grey', alpha =.5 ) + 
  xlab('The average marginal effect') +
  ggtitle('The average marginal effect of Black at the representative ages') +
  theme_bw() +
  xlim(c(0,0.30))
```
