---
title: "Stats 506, F18, Problem Set 4"
author: "Zi Wang, tlwangzi@umich.edu"
date: "12/11/2018"
output: html_document
---
## Question 1

```{r q1_source, echo = FALSE, message = FALSE}
library(tidyverse)
source('./ps4_q1.R')
```

The result:

```{r q1a, echo = FALSE}
cap = '**Table 1.** The all-time leader in hits (with at least 200 hits)
       for each birth country.'
knitr::kable(result, align = 'r', caption = cap)
```

\pagebreak

## Question 2 

```{r q2a_source, echo = FALSE, message = FALSE}
source('./ps4_q2_funcs.R')
source('./ps4_q2a.R')
```

### a.

Show the dimension and the first 20 rows of `results_q4a`:  

```{r q2a, echo = TRUE}
dim(results_q4a)
cap = '**Table 2.** The first 20 rows of the results 
       for Monte Carlo simulations using mclapply.'
knitr::kable(results_q4a[1:20,], align = 'r', digits = 5, caption = cap)
```

### b.
The detail can be seen in `ps4_q2b.R` and `run_ps4_q2b.pbs`.  
The result can be seen in `results_q4b.RData` and `ps4_q2b.Rout`.  
Here, we produce a table to show the dimension and 
first 20 rows of `results_q4b`:  
```{r q2b, echo = TRUE}
load("./results_q4b.RData")
dim(results_q4b)
cap = '**Table 3.** The first 20 rows of the results 
       for Monte Carlo simulations using doParallel
       and nested foreach loops.'
knitr::kable(results_q4b[1:20,], align = 'r', digits = 5, caption = cap)
```

### c.
The detail can be seen in `ps4_q2c.R` and `run_ps4_q2c.pbs`.    
The result can be seen in `ps4_q2c-1.Rout`,
  `ps4_q2c-2.Rout` and `ps4_q2c-4.Rout`.   
Here, we source `ps4_q2c.R`, and produce a table to show the combination of 
the first 10 rows of the result from the situation of sigma = 0.25,
sigma = 0.5 and sigma = 1:  
```{r q2c_source, echo = TRUE, message = FALSE}
source("./ps4_q2c.R")
```

```{r q2c, echo = TRUE}
cap = '**Table 4.** The combination of the 
first 10 rows of the result using future from the situation of sigma = 0.25,
sigma = 0.5 and sigma = 1.'
knitr::kable(results_q4c[c(1:10, 113:122, 225:234),], align = 'r', digits = 5, 
             caption = cap)
```

## Question 3
The detail can be seen in `ps4_q3.sas`.  
Results from (c) can be seen in `ps4_q3c.csv`.  
Results from (d) can be seen in `ps4_q3d.csv`.  
We read data from `ps4_q3c.csv` and `ps4_q3d.csv` and verify they match:
```{r q3_match, echo = TRUE, message = FALSE}
result_c = read_delim("ps4_q3c.csv", delim = ',')
cap = "**Table 5.** Result from c"
knitr::kable(result_c, align = 'r', caption = cap)
result_d = read_delim("ps4_q3d.csv", delim = ',')
cap = "**Table 6.** Result from d"
knitr::kable(result_d, align = 'r', caption = cap)
result_c == result_d
```
Also, we can use `proc compare` in SAS to verify they match, 
and the result shows that they match.
