---
title: "Stats 506, F18, Problem Set 1"
author: "Zi Wang, tlwangzi@umich.edu"
date: "10/02/2018"
output: pdf_document
---
## Question 1

```{bash setup_q1, echo=FALSE}
#!/bin/bash

## Analysis of the RECS 2015 data.
##
## The RECS 2015 Data used in this script can be found at the link below:
##   https://www.eia.gov/consumption/residential/data/2015/csv/recs2015_public_v3.csv
##
## Author: Zi Wang (tlwangzi@umich.edu)
## Updated: Oct 02, 2018
```

### Part A

#### i.
```{bash q1Ai}
cat ./recs2015_public_v3.csv|cut -b 10|grep ^3|wc -l
```

#### ii.
```{bash q1Aii}
cat ./recs2015_public_v3.csv|cut -d , -f 1,479-575 >recs_comp.txt|gzip recs_comp.txt
```


### Part B
#### i.
```{bash q1Bi, echo=FALSE}
for i in {1..4}
do
cat ./recs2015_public_v3.csv|cut -b 10|grep ^$i | wc -l
done
```

#### ii.
```{bash q1Bii, echo=FALSE}
cat ./recs2015_public_v3.csv |cut -d , -f 2,3 |tail -n +2 |sort -n -k 4 -t'"'|uniq   > region_division.txt
cat region_division.txt
```

\pagebreak

## Question 2 

```{r setup_q2, echo = FALSE, include = FALSE}
library(nycflights13)
suppressMessages(library("tidyverse"))
source('ps1_q2.R')
q2 = ps1_q2()
```

### a.

```{r q2a, echo = FALSE, fig.height = 4, fig.cap = cap}
air_name = q2[[1]]
cap = "Airlines responsible for at least 1% of the flights"
cols = c('airlines')
knitr::kable(air_name,digits = 1,caption = cap,col.names = cols)
```

### b.
```{r q2b, echo = FALSE, fig.height = 4, fig.cap = cap}
num_2013 = q2[[2]]
num_2014 = q2[[3]]
num_diff = q2[[4]]
num_diff_max = q2[[5]]
num_diff_min = q2[[6]]
cap = "The number and percent of annual flights 
  in the first 10 months of 2013"
cols = c('carrier','number','percent','lwr of CI','upr of CI')
knitr::kable(num_2013,digits = 1,caption = cap,col.names = cols)
cap = "The number and percent of annual flights 
  in the first 10 months of 2014"
cols=c('carrier','number','percent','lwr of CI','upr of CI')
knitr::kable(num_2014,digits = 1,caption = cap,col.names = cols)
cap = "The change"
cols=c('carrier','change in number','change in percent',
       'lwr of CI','upr of CI')
knitr::kable(num_diff,digits = 1,caption = cap,col.names = cols)
```

The lines showed the largest increase:
```{r increase, echo = FALSE}
num_diff_max
```
The lines showed the largest decrease:
```{r decrease, echo = FALSE}
num_diff_min
```

Reason:The total number of flights in the first 10 months of 2014 
  is less than that of 2013.


### c.
```{r q2c, echo = FALSE, fig.height = 4, fig.cap = cap}
per_tol = q2[[7]]
per_tol_JFK_max = q2[[8]]
per_tol_EWR_max = q2[[9]]
per_tol_LGA_max = q2[[10]]
cap = "The percent of flights each airline is responsible for 
  among of the three NYC airports."
cols=c('carrier','airport','percent','lwr of CI','upr of CI')
knitr::kable(per_tol,digits = 1,caption = cap,col.names = cols)
```

The largest carrier at airport JFK:
```{r JFK, echo = FALSE}
per_tol_JFK_max
```
The largest carrier at airport EWR:
```{r EWR, echo = FALSE}
per_tol_EWR_max
```
The largest carrier at airport LGA.
```{r LGA, echo = FALSE}
per_tol_LGA_max
```

\pagebreak
## Question 3

``` {r setup_q3, echo = FALSE, include = FALSE}
source('ps1_q3.R')
q3=ps1_q3()
```

### a. 
```{r q3a, echo = FALSE, fig.height = 4, fig.cap = cap}
recs_wall_summary = q3[[1]]
recs_wall_max = q3[[2]]
recs_wall_min = q3[[3]]
cap = "Percent of homes which have stucco construction 
  as the major outside wall material"
cols = c('Division','Percent','Standard errors')
knitr::kable(recs_wall_summary,digits = 1,caption = cap,col.names = cols)
```

The division which has the largest proportion:
```{r largest, echo = FALSE}
recs_wall_max
```
The division which has the lowest proportion:
```{r lowest, echo = FALSE}
recs_wall_min
```


### b.
```{r q3b, echo = FALSE, fig.height = 4, fig.cap = cap}
recs_elec_summary = q3[[4]]
recs_elec_summary_stra = q3[[5]]
cap = "Average total electricity usage in kilowatt hours in each division"
cols=c('Division','Average','Standard errors')
knitr::kable(recs_elec_summary,digits = 1,caption = cap,col.names = cols)
cap = "Average total electricity usage in kilowatt hours in each division 
  stratified by urban and rural status"
cols=c('Division','Status','Average','Standard errors')
knitr::kable(recs_elec_summary_stra,digits = 1,caption = cap,col.names = cols)
```


### c.
```{r q3c, echo = FALSE, fig.height = 4, fig.cap = cap}
recs_int_summary_U = q3[[6]]
recs_int_summary_R = q3[[7]]
rec_int_diff_max = q3[[8]]
cap = "The proportion of homes with internet access in urban areas"
cols=c('Division','Proportion','Standard errors')
knitr::kable(recs_int_summary_U,digits = 1,caption = cap,col.names = cols)
cap = "The proportion of homes with internet access in rural areas"
cols=c('Division','Proportion','Standard errors')
knitr::kable(recs_int_summary_R,digits = 1,caption = cap,col.names = cols)
```

The division which has the largest disparity:
```{r disparity, echo = FALSE}
rec_int_diff_max
```
